---
title: "Time Series Analysis & Forecasting Using R"
subtitle: "4. Seasonality and trends"
---

## Outline

\vspace*{0.7cm}\tableofcontents


```{r}
#| label: setup
#| include: false
#| cache: false
source("setup.R")
library(purrr)
library(transformr) # Just to get it on renv
library(gganimate)
library(latex2exp)
```

# Time series decompositions

## Time series decomposition

Trend-Cycle
: aperiodic changes in level over time.

Seasonal
: (almost) periodic changes in level due to seasonal factors (e.g., the quarter of the year, the month, or day of the week).

\begin{block}{Additive decomposition}\vspace*{-0.3cm}
\[ y_t = S_t + T_t + R_t \]
\end{block}
\begin{tabular}{@{}llp{8cm}@{}}
where & $y_t=$ & data at period $t$ \\
      & $T_t=$ & trend-cycle component at period $t$\\
      & $S_t=$ & seasonal component at period $t$ \\
      & $R_t=$ & remainder component at period $t$
\end{tabular}

## STL decomposition

\fontsize{13}{14}\sf

  *  STL: "Seasonal and Trend decomposition using Loess"
  *  Very versatile and robust.
  *  Seasonal component allowed to change over time, and rate of change controlled by user.
  *  Smoothness of trend-cycle also controlled by user.
  *  Optionally robust to outliers
  *  No trading day or calendar adjustments.
  *  Only additive.
  *  Take logs to get multiplicative decomposition.
  *  Use Box-Cox transformations to get other decompositions.

## US Retail Employment
\fontsize{11}{11}\sf

```{r usretail}
us_retail_employment <- us_employment |>
  filter(year(Month) >= 1990, Title == "Retail Trade") |>
  select(-Series_ID)
us_retail_employment
```

\vspace*{10cm}

## US Retail Employment

```{r dable1}
us_retail_employment |>
  autoplot(Employed) +
  labs(y = "Persons (thousands)", title = "Total employment in US retail")
```

\vspace*{10cm}

## US Retail Employment
\fontsize{11}{11}\sf

```{r dable2}
dcmp <- us_retail_employment |>
  model(stl = STL(Employed))
dcmp
```

\vspace*{10cm}

## US Retail Employment
\fontsize{11}{11}\sf

```{r dable3}
components(dcmp)
```

\vspace*{10cm}

## US Retail Employment

```{r usretail-stl, fig.width=8, fig.height=4}
components(dcmp) |> autoplot()
```

## US Retail Employment

```{r dable4, fig.height=2.7}
us_retail_employment |>
  autoplot(Employed, color = "gray") +
  autolayer(components(dcmp), trend, color = "#D55E00") +
  labs(y = "Persons (thousands)", title = "Total employment in US retail")
```

\vspace*{10cm}

## US Retail Employment

```{r usretail3}
components(dcmp) |> gg_subseries(season_year)
```

## Seasonal adjustment

  *  Useful by-product of decomposition:  an easy way to calculate seasonally adjusted data.
  *  Additive decomposition: seasonally adjusted data given by
$$y_t - S_t = T_t + R_t$$
  *  Multiplicative decomposition: seasonally adjusted data given by
$$y_t / S_t = T_t \times R_t$$

## US Retail Employment

```{r usretail-sa, fig.height=2.7}
us_retail_employment |>
  autoplot(Employed, color = "gray") +
  autolayer(components(dcmp), season_adjust, color = "#0072B2") +
  labs(y = "Persons (thousands)", title = "Total employment in US retail")
```

## Seasonal adjustment

  * We use estimates of $S$ based on past values to seasonally adjust a current value.
  *  Seasonally adjusted series reflect **remainders** as well as **trend**. Therefore they are not "smooth" and "downturns" or "upturns" can be misleading.
  *  It is better to use the trend-cycle component to look for turning points.

## STL decomposition

```{r stlwindowanim, echo=FALSE, warning=FALSE, message=FALSE, fig.show='animate', interval=1/10,  fig.height=5.35, fig.width=8, aniopts='controls,buttonsize=0.3cm,width=11.5cm', eval=TRUE}
s_windows <- seq(5, 55, by = 2)
stl_defs <- purrr::map(s_windows, function(s_window) {
  STL(Employed ~ season(window = s_window), robust = TRUE)
})
names(stl_defs) <- sprintf("season(window=%02d)", s_windows)

us_retail_employment |>
  model(!!!stl_defs) |>
  components() |>
  as_tibble() |>
  pivot_longer(Employed:remainder,
    names_to = "component", names_ptypes = list(component = factor(levels = c("Employed", "trend", "season_year", "remainder"))),
    values_to = "Employed"
  ) |>
  ggplot(aes(x = Month, y = Employed)) +
  geom_line() +
  facet_grid(rows = vars(component), scales = "free_y") +
  labs(
    title = "STL decomposition of US retail employment",
    subtitle = "{closest_state}"
  ) +
  transition_states(.model)
```

\vspace*{10cm}

## STL decomposition

```{r echo = TRUE, results = 'hide'}
us_retail_employment |>
  model(STL(Employed ~ trend(window = 15) + season(window = "periodic"),
    robust = TRUE
  )) |>
  components()
```

\fontsize{12}{13}\sf

  *  `trend(window = ?)` controls wiggliness of trend component.
  *  `season(window = ?)` controls variation on seasonal component.
  *  `season(window = 'periodic')` is equivalent to an infinite window.

## STL decomposition

```{r mstl, fig.width=8.5, fig.height=3.4}
us_retail_employment |>
  model(STL(Employed)) |>
  components() |>
  autoplot()
```

\only<2>{\begin{textblock}{7}(8,0.2)\fontsize{11}{11}\sf
\begin{alertblock}{}
\begin{itemize}\tightlist
\item \texttt{STL()} chooses \texttt{season(window=13)} by default
\item Can include transformations.
\end{itemize}
\end{alertblock}
\end{textblock}}

## STL decomposition
\fontsize{13}{15}\sf

* Algorithm that updates trend and seasonal components iteratively.
* Starts with $\hat{T}_t=0$
* Uses a mixture of loess and moving averages to successively refine the trend and seasonal estimates.
* trend window controls loess bandwidth on deasonalised values.
* season window controls loess bandwidth on detrended subseries.
* Robustness weights based on remainder.
* Default season: `window = 13`
* Default trend:\newline\mbox{}\hfill `window = nextodd(ceiling((1.5*period)/(1-(1.5/s.window)))`
* `window` values should be odd numbers for symmetry.

## Australian holidays

```{r holidays, include=FALSE}
holidays <- tourism |>
  filter(Purpose == "Holiday") |>
  group_by(State) |>
  summarise(Trips = sum(Trips))
```

```{r holidays-plot2, echo=TRUE, dependson="holidays", fig.height=2.8}
holidays |> autoplot(Trips) +
  labs(y = "thousands of trips", x = "Year",
       title = "Australian domestic holiday nights")
```

## Australian holidays

```{r stlagain2, echo=TRUE, warning=FALSE, fig.width=8, fig.height=3}
holidays |>
  model(stl = STL(Trips)) |>
  components() |>
  autoplot()
```

## Holidays decomposition
\fontsize{9}{10}\sf

```{r holidays2}
dcmp <- holidays |>
  model(stl = STL(Trips)) |>
  components()
dcmp
```

## Holidays decomposition
\fontsize{9}{10}\sf

```{r holidays3, fig.height=4.6}
dcmp |> gg_subseries(season_year)
```

## Holidays decomposition
\fontsize{9}{10}\sf

```{r holidays-trend, message=FALSE, warning=FALSE, fig.height=3.5, out.height="70%"}
autoplot(dcmp, trend, scale_bars = FALSE) +
  autolayer(holidays, alpha = 0.4)
```

# Lab Session 8

## Lab Session 8
\fontsize{13}{14}\sf

1. Produce the following decomposition

    \fontsize{9}{11}\sf

    ```r
    canadian_gas |>
      model(STL(Volume ~ season(window=7) + trend(window=11))) |>
      components() |>
      autoplot()
    ```

    \fontsize{13}{14}\sf

2. What happens as you change the values of the two `window` arguments?

3. How does the seasonal shape change over time? [Hint: Try plotting the seasonal component using `gg_season`.]

4. Can you produce a plausible seasonally adjusted series? [Hint: `season_adjust` is one of the variables returned by `STL`.]

# Multiple seasonality

## Multiple seasonality
\fontsize{9}{9}\sf

```r
vic_elec |>
  model(STL(Demand)) |>
  components() |>
  autoplot()
```

```{r vic_elec, echo=FALSE}
png("figs/vic_elec_stl.png", width = 18, height = 18, res = 300, units = "cm", type = "cairo-png")
vic_elec |>
  model(STL(Demand)) |>
  components() |>
  autoplot()
crop::dev.off.crop()
```

\placefig{7}{0.2}{width=8.7cm, height=20cm}{vic_elec_stl}

\vspace*{10cm}

# The ABS stuff-up

## The ABS stuff-up

\full{abs1}

## The ABS stuff-up

\full{abs2}

## The ABS stuff-up

\full{abs3}

## The ABS stuff-up

```{r abs1, echo=FALSE}
employed <- tsibble(
  Time = yearmonth("1978 Feb") + 0:439,
  Employed = c(
    5985.7, 6040.6, 6054.2, 6038.3, 6031.3, 6036.1, 6005.4, 6024.3, 6045.9, 6033.8, 6125.4, 5971.3,
    6050.7, 6096.2, 6087.7, 6075.6, 6095.7, 6103.9, 6078.5, 6157.8, 6164.0, 6188.8, 6257.2, 6112.9,
    6207.2, 6278.7, 6224.9, 6273.4, 6269.9, 6314.1, 6281.4, 6360.0, 6320.2, 6342.0, 6426.6, 6253.0,
    6356.5, 6428.1, 6426.3, 6412.4, 6413.9, 6425.3, 6393.7, 6502.7, 6445.3, 6433.3, 6506.9, 6355.5,
    6432.4, 6497.4, 6431.6, 6440.9, 6414.3, 6425.9, 6379.3, 6443.5, 6421.1, 6366.8, 6370.1, 6172.0,
    6263.9, 6310.3, 6254.5, 6272.8, 6266.5, 6295.0, 6241.2, 6358.2, 6336.1, 6377.5, 6456.5, 6251.4,
    6365.4, 6503.2, 6477.6, 6489.7, 6499.0, 6528.7, 6466.1, 6579.8, 6553.2, 6576.1, 6636.0, 6452.4,
    6595.7, 6657.4, 6588.8, 6657.9, 6659.4, 6703.4, 6675.5, 6814.7, 6771.1, 6881.9, 6910.8, 6753.6,
    6861.9, 6961.9, 6997.9, 6979.0, 7007.7, 6991.5, 6918.5, 7040.6, 7030.4, 7034.2, 7116.8, 6902.5,
    7022.3, 7133.4, 7109.6, 7103.5, 7128.9, 7175.6, 7092.3, 7186.5, 7177.4, 7182.2, 7330.7, 7169.4,
    7247.3, 7397.4, 7383.4, 7354.8, 7378.3, 7383.1, 7353.3, 7503.2, 7477.3, 7508.6, 7622.9, 7423.8,
    7566.5, 7634.6, 7678.4, 7720.8, 7711.0, 7740.8, 7715.3, 7841.6, 7806.5, 7862.4, 7935.5, 7707.7,
    7803.0, 7874.1, 7887.9, 7908.5, 7900.3, 7919.4, 7808.0, 7905.5, 7848.9, 7826.9, 7915.5, 7641.3,
    7708.7, 7715.4, 7717.2, 7703.7, 7678.1, 7583.0, 7620.7, 7713.2, 7638.0, 7614.9, 7712.2, 7518.9,
    7597.2, 7646.2, 7644.1, 7631.4, 7637.3, 7668.3, 7613.4, 7709.7, 7665.7, 7587.4, 7693.4, 7533.7,
    7531.0, 7645.7, 7572.6, 7620.5, 7627.9, 7646.5, 7589.4, 7747.6, 7738.8, 7744.9, 7842.1, 7646.8,
    7738.6, 7824.2, 7827.4, 7857.9, 7878.4, 7966.0, 7861.7, 8054.4, 7997.2, 8003.3, 8135.5, 7928.4,
    8049.9, 8118.1, 8174.6, 8165.2, 8205.6, 8229.0, 8165.9, 8300.4, 8232.6, 8300.3, 8395.7, 8166.7,
    8246.6, 8280.4, 8248.0, 8297.1, 8311.7, 8332.1, 8265.9, 8373.0, 8319.4, 8314.4, 8431.4, 8235.2,
    8291.4, 8347.5, 8343.1, 8330.2, 8345.6, 8374.9, 8250.3, 8474.0, 8405.2, 8462.1, 8540.5, 8334.7,
    8413.0, 8460.0, 8499.9, 8482.5, 8516.8, 8541.9, 8455.2, 8653.2, 8601.0, 8554.3, 8696.5, 8477.4,
    8556.5, 8618.9, 8631.9, 8606.5, 8673.2, 8706.7, 8603.6, 8777.1, 8755.3, 8763.7, 8900.7, 8628.2,
    8754.4, 8830.7, 8882.2, 8865.0, 8922.0, 9020.0, 8911.6, 9061.3, 8973.1, 8912.7, 9059.6, 8834.9,
    8920.9, 8956.0, 9023.6, 9004.6, 9021.9, 9048.9, 8971.9, 9105.9, 9058.7, 9055.6, 9177.1, 8993.4,
    9092.3, 9128.5, 9129.5, 9134.7, 9180.8, 9194.5, 9150.3, 9303.5, 9249.1, 9286.7, 9439.7, 9281.7,
    9372.6, 9362.1, 9365.6, 9380.1, 9370.4, 9363.9, 9327.0, 9486.1, 9447.8, 9427.7, 9573.6, 9363.8,
    9434.5, 9506.4, 9512.0, 9533.5, 9543.3, 9553.1, 9462.1, 9668.6, 9662.2, 9652.9, 9807.8, 9634.4,
    9744.6, 9828.3, 9856.3, 9850.8, 9896.6, 9912.3, 9870.3, 10004.6, 9949.7, 9945.0, 10074.7, 9842.7,
    9961.1, 10048.7, 10041.0, 10082.1, 10120.8, 10170.8, 10105.8, 10299.5, 10212.4, 10201.6, 10404.3,
    10156.1, 10277.0, 10349.2, 10362.9, 10412.0, 10436.3, 10456.8, 10406.4, 10588.8, 10520.5, 10535.0,
    10710.1, 10524.9, 10622.9, 10677.4, 10706.2, 10690.3, 10745.0, 10761.9, 10710.4, 10854.5, 10807.4,
    10757.3, 10915.6, 10681.0, 10776.7, 10775.2, 10792.7, 10786.8, 10770.9, 10808.8, 10707.3, 10882.1,
    10845.2, 10829.2, 11010.9, 10809.9, 10889.2, 10928.9, 10940.1, 10957.4, 11009.3, 11030.5, 10973.8,
    11159.4, 11129.0, 11144.5, 11295.0, 11063.7, 11146.2, 11217.0, 11186.5, 11196.2, 11221.3, 11227.5,
    11130.7, 11321.2, 11274.0, 11240.6, 11354.8, 11159.0, 11236.2, 11332.4, 11328.3, 11389.0, 11350.6,
    11363.7, 11259.8, 11452.6, 11401.9, 11375.0, 11518.4, 11304.0, 11424.3, 11436.3, 11482.2, 11495.6,
    11497.8, 11486, 11369, 11547, 11499, 11472, 11571, 11354, 11493, 11562, 11589, 11595, 11602, 11590,
    11622, 11593
  ),
  index = Time
) |>
  mutate(
    Month = month(Time, label = TRUE),
    Year = year(Time)
  ) |>
  select(Time, Month, Year, Employed)
```

\fontsize{11}{11}\sf

```{r abs2}
employed
```

## The ABS stuff-up

```{r abs3}
employed |>
  autoplot(Employed) +
  labs(title = "Total employed", y = "Thousands")
```

## The ABS stuff-up

```{r abs4, fig.height=2.8}
employed |>
  filter(Year >= 2005) |>
  autoplot(Employed) +
  labs(title = "Total employed", y = "Thousands")
```

## The ABS stuff-up

```{r abs5, fig.height=2.8}
employed |>
  filter(Year >= 2005) |>
  gg_season(Employed, labels = "right") +
  labs(title = "Total employed", y = "Thousands")
```

## The ABS stuff-up

```{r abs6, fig.height=2}
employed |>
  mutate(diff = difference(Employed)) |>
  filter(Month == "Sep") |>
  ggplot(aes(y = diff, x = 1)) +
  geom_boxplot() +
  coord_flip() +
  labs(title = "Sep - Aug: total employed", y = "Thousands") +
  scale_x_continuous(breaks = NULL, labels = NULL)
```

## The ABS stuff-up

```{r abs7, fig.height=2.8}
dcmp <- employed |>
  filter(Year >= 2005) |>
  model(stl = STL(Employed ~ season(window = 11), robust = TRUE))
components(dcmp) |> autoplot()
```

## The ABS stuff-up

```{r abs8, fig.height=2.8}
components(dcmp) |>
  filter(year(Time) == 2013) |>
  gg_season(season_year) +
  labs(title = "Seasonal component") + guides(colour = "none")
```

## The ABS stuff-up

```{r abs9}
components(dcmp) |>
  as_tsibble() |>
  autoplot(season_adjust)
```

## The ABS stuff-up
\fontsize{13}{15}\sf

  *  August 2014 employment numbers higher than expected.
  *  Supplementary survey usually conducted in August for employed people.
  *  Most likely, some employed people were claiming to be unemployed in August to avoid supplementary questions.
  *  Supplementary survey not run in 2014, so no motivation to lie about employment.
  *  In previous years, seasonal adjustment fixed the problem.
  *  The ABS has now adopted a new method to avoid the bias.
