---
title: "Time Series Analysis & Forecasting Using R"
subtitle: "5. Time series features"
---

## Outline

\vspace*{0.7cm}\tableofcontents

```{r}
#| label: setup
#| include: false
#| cache: false
source("setup.R")
library(glue)
library(stringr)
```

# STL Features

## Strength of seasonality and trend

\begin{alertblock}{STL decomposition}
\centerline{$y_t = T_t+S_t+R_t$}
\end{alertblock}

### Seasonal strength
$$\max\left(0, 1-\frac{\text{Var}(R_t)}{\text{Var}(S_t+R_t)}\right)$$

### Trend strength
$$\max\left(0, 1-\frac{\text{Var}(R_t)}{\text{Var}(T_t+R_t)}\right)$$

## Feature extraction and statistics

\fontsize{9}{10}\sf

```{r features}
tourism |> features(Trips, feat_stl)
```

## Feature extraction and statistics
\fontsize{8}{9}\sf

```{r}
#| label: features-plot
#| fig-height: 3.6
tourism |>
  features(Trips, feat_stl) |>
  ggplot(aes(x = trend_strength, y = seasonal_strength_year, col = Purpose)) +
  geom_point() + facet_wrap(vars(State))
```

\only<2->{\begin{textblock}{4.7}(9.8,6.9)
\begin{alertblock}{}\fontsize{10}{10}\sf
\begin{itemize}\tightlist
\item Holidays more seasonal than other travel.
\item WA has strongest trends.
\end{itemize}
\end{alertblock}\end{textblock}}

## Feature extraction and statistics
\fontsize{9}{9}\sf

Find the most seasonal time series:

```{r extreme}
most_seasonal <- tourism |>
  features(Trips, feat_stl) |>
  filter(seasonal_strength_year == max(seasonal_strength_year))
```

\pause\vspace*{-0.3cm}

```{r extreme2, fig.height=1.8}
tourism |>
  right_join(most_seasonal, by = c("State", "Region", "Purpose")) |>
  ggplot(aes(x = Quarter, y = Trips)) +
  geom_line() + facet_grid(vars(State, Region, Purpose))
```

## Feature extraction and statistics
\fontsize{9}{9}\sf

Find the most trended time series:

```{r extreme3}
most_trended <- tourism |>
  features(Trips, feat_stl) |>
  filter(trend_strength == max(trend_strength))
```

\pause\vspace*{-0.3cm}

```{r extreme4, fig.height=1.8}
tourism |>
  right_join(most_trended, by = c("State", "Region", "Purpose")) |>
  ggplot(aes(x = Quarter, y = Trips)) +
  geom_line() + facet_grid(vars(State, Region, Purpose))
```

# Lab Session 9

## Lab Session 9

 * Use ``GGally::ggpairs()`` to look at the relationships between the STL-based features. You might wish to change `seasonal_peak_year` and `seasonal_trough_year` to factors.
 * Which is the peak quarter for holidays in each state?

## Feature extraction and statistics

\fontsize{9}{10}\sf

```{r featuresacf}
tourism |> features(Trips, feat_acf)
```

# Dimension reduction for features

## Feature extraction and statistics

```{r pca}
#| echo: false
#| include: false
# Save pdf figures
savepdf <- function(file, width = 16, height = 10) {
  fname <<- paste("figs/", file, ".pdf", sep = "")
  Cairo::CairoPDF(fname, width = width / 2.54, height = height / 2.54, pointsize = 10)
  par(mgp = c(2.2, 0.45, 0), tcl = -0.4, mar = c(3.3, 3.6, 1.1, 1.1))
}
endpdf <- function() {
  crop::dev.off.crop(fname)
}
# Compute features
tourism_features <- tourism |>
  features(Trips, feature_set(pkgs = "feasts"))
# Compute PCs
pcs <- tourism_features |>
  select(-State, -Region, -Purpose) |>
  prcomp(scale = TRUE) |>
  broom::augment(tourism_features)
# Save some PC plots
savepdf("pca1", 18, 10)
pcs |> ggplot(aes(x = .fittedPC1, y = .fittedPC2)) +
  geom_point() +
  theme(aspect.ratio = 1)
endpdf()
savepdf("pca2", 18, 10)
pcs |> ggplot(aes(x = .fittedPC1, y = .fittedPC2, col = State)) +
  geom_point() +
  theme(aspect.ratio = 1)
endpdf()
savepdf("pca3", 18, 10)
pcs |> ggplot(aes(x = .fittedPC1, y = .fittedPC2, col = Purpose)) +
  geom_point() +
  theme(aspect.ratio = 1)
endpdf()
# Find outliers
outliers <- pcs |>
  filter(.fittedPC1 > 12 | (.fittedPC1 > 10 & .fittedPC2 > 0))
savepdf("pca4", 18, 10)
pcs |> ggplot(aes(x = .fittedPC1, y = .fittedPC2, col = Purpose)) +
  geom_point() +
  theme(aspect.ratio = 1) +
  geom_point(data = outliers, aes(x = .fittedPC1, y = .fittedPC2), col = "black", shape = 1, size = 3)
endpdf()
```

```{r}
#| label: tourismfeatures
#| eval: false
tourism_features <- tourism |>
  features(Trips, feature_set(pkgs = "feasts"))
```

\fontsize{8}{8}\sf\vspace*{-0.3cm}

```{r}
#| echo: false
tourism_features
```

\begin{textblock}{4.5}(10.6,.5)
\begin{alertblock}{}\fontsize{10}{12}\sf
All features from the feasts package
\end{alertblock}
\end{textblock}

## Feature extraction and statistics

```{r pcatable}
pcs <- tourism_features |>
  select(-State, -Region, -Purpose) |>
  prcomp(scale = TRUE) |>
  broom::augment(tourism_features)
```

\fontsize{8}{8}\sf\vspace*{-0.3cm}

```{r echo=FALSE}
pcs
```

\begin{textblock}{4.5}(10.6,.5)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

## Feature extraction and statistics
\fontsize{9}{9}\sf

\begin{textblock}{3.3}(1.4,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

```r
pcs |> ggplot(aes(x=.fittedPC1, y=.fittedPC2)) +
  geom_point() + theme(aspect.ratio=1)
```

\placefig{7}{2.6}{height=6.4cm, width=12cm}{pca1}
\vspace*{10cm}

## Feature extraction and statistics
\fontsize{9}{9}\sf

\begin{textblock}{3.3}(1.4,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

```r
pcs |> ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=State)) +
  geom_point() + theme(aspect.ratio=1)
```

\placefig{7}{2.6}{height=6.4cm, width=12cm}{pca2}
\vspace*{10cm}

## Feature extraction and statistics
\fontsize{9}{9}\sf

\begin{textblock}{3.3}(1.4,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

```r
pcs |> ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=Purpose)) +
  geom_point() + theme(aspect.ratio=1)
```

\only<1>{\placefig{7}{2.6}{height=6.4cm, width=12cm}{pca3}}
\only<2>{\placefig{7}{2.6}{height=6.4cm, width=12cm}{pca4}}
\vspace*{10cm}

## Feature extraction and statistics
\fontsize{8}{8}\sf

```{r outliers2, fig.height=4.5, fig.width=12, out.height="55%"}
outliers |>
  left_join(tourism, by = c("State", "Region", "Purpose")) |>
  mutate(Series = glue("{State}", "{Region}", "{Purpose}", .sep = "\n\n")) |>
  ggplot(aes(x = Quarter, y = Trips)) +
  geom_line() + facet_grid(Series ~ .) +
  labs(title = "Outlying time series in PC space")
```

# Lab Session 10

## Lab Session 10

* Use a feature-based approach to look for outlying series in `PBS`.

* What is unusual about the series you identify as outliers?
