---
title: "Time Series Analysis & Forecasting Using R"
subtitle: "8. ARIMA models"
---

## Outline

\vspace*{0.7cm}\tableofcontents


```{r}
#| label: setup
#| include: false
#| cache: false
source("setup.R")
```

# ARIMA models

## ARIMA models

\begin{tabular}{rl}
\textbf{AR}: & autoregressive (lagged observations as inputs)\\
\textbf{I}: & integrated (differencing to make series stationary)\\
\textbf{MA}: & moving average (lagged errors as inputs)
\end{tabular}

\pause

###
An ARIMA model is rarely interpretable in terms of visible data structures like trend and seasonality. But it can capture a huge range of time series patterns.

## Stationarity

\begin{block}{Definition}
If $\{y_t\}$ is a stationary time series, then for all $s$, the distribution of $(y_t,\dots,y_{t+s})$ does not depend on $t$.
\end{block}\pause

A **stationary series** is:

* roughly horizontal
* constant variance
* no patterns predictable in the long-term

## Stationary?
\fontsize{11}{12}\sf

```{r, fig.height=2.7}
gafa_stock |>
  filter(Symbol == "GOOG", year(Date) == 2018) |>
  autoplot(Close) +
  labs(y = "Google closing stock price ($US)")
```

## Stationary?
\fontsize{11}{12}\sf

```{r, fig.height=2.7}
gafa_stock |>
  filter(Symbol == "GOOG", year(Date) == 2018) |>
  autoplot(difference(Close)) +
  labs(y = "Daily change in Google closing stock price")
```

## Differencing
\fontsize{13}{15}\sf

* Differencing helps to **stabilize the mean**.
* The differenced series is the *change* between each observation in the original series.
* Occasionally the differenced data will not appear stationary and it may be necessary to difference the data a second time.
* In practice, it is almost never necessary to go beyond second-order differences.

## Autoregressive models

\begin{block}{Autoregressive (AR) models:}\vspace*{-0.3cm}
$$
  y_{t} = c + \phi_{1}y_{t - 1} + \phi_{2}y_{t - 2} + \cdots + \phi_{p}y_{t - p} + \varepsilon_{t},
$$
where $\varepsilon_t$ is white noise. A multiple regression with \textbf{lagged values} of $y_t$ as predictors.
\end{block}

```{r arp, echo=FALSE, fig.height=2.3}
set.seed(1)
p1 <- tsibble(idx = seq_len(100), sim = 10 + arima.sim(list(ar = -0.8), n = 100), index = idx) |>
  autoplot(sim) + labs(x = "time", y = "", title = "AR(1)")
p2 <- tsibble(idx = seq_len(100), sim = 20 + arima.sim(list(ar = c(1.3, -0.7)), n = 100), index = idx) |>
  autoplot(sim) + labs(x = "time", y = "", title = "AR(2)")
gridExtra::grid.arrange(p1, p2, nrow = 1)
```

\vspace*{-0.4cm}

* Cyclic behaviour is possible when $p\ge 2$.

## Moving Average (MA) models

\begin{block}{Moving Average (MA) models:}\vspace*{-0.3cm}
$$
  y_{t} = c + \varepsilon_t + \theta_{1}\varepsilon_{t - 1} + \theta_{2}\varepsilon_{t - 2} + \cdots + \theta_{q}\varepsilon_{t - q},
$$
where $\varepsilon_t$ is white noise.
A multiple regression with \textbf{lagged \emph{errors}} as predictors. \emph{Don't confuse with moving average smoothing!}
\end{block}

```{r maq, fig.height=2.3, echo=FALSE}
set.seed(2)
p1 <- tsibble(idx = seq_len(100), sim = 20 + arima.sim(list(ma = 0.8), n = 100), index = idx) |>
  autoplot(sim) + labs(x = "time", y = "", title = "MA(1)")
p2 <- tsibble(idx = seq_len(100), sim = arima.sim(list(ma = c(-1, +0.8)), n = 100), index = idx) |>
  autoplot(sim) + labs(x = "time", y = "", title = "MA(2)")

gridExtra::grid.arrange(p1, p2, nrow = 1)
```

## ARIMA models

\begin{block}{Autoregressive Moving Average models:}\vspace*{-0.7cm}
\begin{align*}
  y_{t} &= c + \phi_{1}y_{t - 1} + \cdots + \phi_{p}y_{t - p} \\
        & \hspace*{2.4cm}\text{} + \theta_{1}\varepsilon_{t - 1} + \cdots + \theta_{q}\varepsilon_{t - q} + \varepsilon_{t}.
\end{align*}
\end{block}\pause

* Predictors include both **lagged values of $y_t$ and lagged errors.**
\pause

### Autoregressive Integrated Moving Average models
* Combine ARMA model with **differencing**.
* $d$-differenced series follows an ARMA model.
* Need to choose $p$, $d$, $q$ and whether or not to include $c$.

## ARIMA models

\begin{block}{ARIMA($p, d, q$) model}
\begin{tabular}{rl}
AR:& $p =$ order of the autoregressive part\\
I: & $d =$ degree of first differencing involved\\
MA:& $q =$ order of the moving average part.
\end{tabular}
\end{block}

* White noise model: ARIMA(0,0,0)
* Random walk: ARIMA(0,1,0) with no constant
* Random walk with drift: ARIMA(0,1,0) with \rlap{const.}
* AR($p$): ARIMA($p$,0,0)
* MA($q$): ARIMA(0,0,$q$)

## Example: National populations
\fontsize{11}{12}\sf

```{r popfit2, echo=TRUE}
fit <- global_economy |>
  model(arima = ARIMA(Population))
fit
```

## Example: National populations
\fontsize{11}{12}\sf

```{r popfit3, echo=TRUE}
fit |>
  filter(Country == "Australia") |>
  report()
```

\only<2>{\begin{textblock}{6.4}(6,4.6)
\begin{alertblock}{}\fontsize{12}{13}\sf
\centerline{$y_t = 2y_{t-1} - y_{t-2} - 0.7 \varepsilon_{t-1} + \varepsilon_t$}
\mbox{}\hfill$\varepsilon_t \sim \text{NID}(0,4\times10^9)$
\end{alertblock}
\end{textblock}}
\vspace*{3cm}

## Understanding ARIMA models

* If $c=0$ and $d=0$, the long-term forecasts will go to zero.
* If $c=0$ and $d=1$, the long-term forecasts will go to a non-zero constant.
* If $c=0$ and $d=2$, the long-term forecasts will follow a straight line.
* If $c\ne0$ and $d=0$, the long-term forecasts will go to the mean of the data.
* If $c\ne0$ and $d=1$, the long-term forecasts will follow a straight line.
* If $c\ne0$ and $d=2$, the long-term forecasts will follow a quadratic trend.

## Understanding ARIMA models
\fontsize{14}{15.5}\sf

### Forecast variance and $d$
  * The higher the value of $d$, the more rapidly the prediction intervals increase in size.
  * For $d=0$, the long-term forecast standard deviation will go to the standard deviation of the historical data.

## Example: National populations
\fontsize{9}{9}\sf

```{r popfc2, echo=TRUE, fig.height=3.4}
fit |>
  forecast(h = 10) |>
  filter(Country == "Australia") |>
  autoplot(global_economy)
```

## How does ARIMA() work?

\begin{alertblock}{Hyndman and Khandakar (JSS, 2008) algorithm:}
\begin{itemize}\tightlist
\item Select no.\ differences $d$ via KPSS test.
\item Select $p$, $q$ and inclusion of $c$ by minimising AICc.
\item Use stepwise search to traverse model space.
\end{itemize}
\end{alertblock}\pause

\begin{block}{}
\centerline{$\displaystyle \text{AICc} = -2 \log(L) + 2(p+q+k+1)\left[1 + \frac{(p+q+k+2)}{T-p-q-k-2}\right]$}
where $L$ is the maximised likelihood fitted to the \textit{differenced} data,
$k=1$ if $c\neq 0$ and $k=0$ otherwise.\pause
\end{block}

Note: Can't compare AICc for different values of $d$.

## How does ARIMA() work?
\fontsize{12.5}{14.5}\sf

Step1:
: Select current model (with smallest AICc) from:\newline
ARIMA$(2,d,2)$\newline
ARIMA$(0,d,0)$\newline
ARIMA$(1,d,0)$\newline
ARIMA$(0,d,1)$
\pause\vspace*{-0.1cm}

Step 2:
: Consider variations of current model:

    * vary one of $p,q,$ from current model by $\pm1$;
    * $p,q$ both vary from current model by $\pm1$;
    * Include/exclude $c$ from current model.

  Model with lowest AICc becomes current model.

\pause\alert{Repeat Step 2 until no lower AICc can be found.}

## How does ARIMA() work?

```{r ARMAgridsearch, echo=FALSE, message=FALSE, warning=FALSE, fig.asp=1, out.width="60%", fig.width=4, fig.height=4}
start <- tribble(
  ~p, ~q,
  0, 0,
  1, 0,
  0, 1,
  2, 2
)
selected <- tribble(
  ~p, ~q,
  2, 2,
  3, 3,
  4, 2
)
griddf <- expand.grid(p = 0:6, q = 0:6) |>
  as_tibble() |>
  left_join(start |> mutate(start = TRUE)) |>
  left_join(selected |> mutate(chosen = TRUE)) |>
  replace_na(list(start = FALSE, chosen = FALSE)) |>
  mutate(
    step = case_when(
      start ~ 1,
      (p - selected$p[1])^2 + (q - selected$q[1])^2 <= 2 ~ 2,
      (p - selected$p[2])^2 + (q - selected$q[2])^2 <= 2 ~ 3,
      (p - selected$p[3])^2 + (q - selected$q[3])^2 <= 2 ~ 4,
      TRUE ~ NA_real_
    )
  ) |>
  left_join(selected |>
    mutate(step = row_number() + 1) |>
    rename(fromp = p, fromq = q)) |>
  mutate(step = as.character(step))
griddf |>
  ggplot(aes(x = q, y = p)) +
  geom_point(aes(alpha = 0.2), colour = "gray", size = 5, shape = 19) +
  geom_segment(aes(x = fromq, y = fromp, xend = q, yend = p, col = step),
    data = griddf |> filter(step %in% "1"),
    arrow = arrow(length = unit(0.15, "inches"), type = "open"),
    size = 1, lineend = "butt"
  ) +
  geom_point(aes(col = step),
    size = 5, shape = 19,
    data = griddf |> filter(step == "1")
  ) +
  geom_point(data = filter(griddf |> filter(step == "1"), chosen), size = 12, shape = 21, stroke = 1.4) +
  scale_y_reverse(breaks = 0:6) +
  scale_x_continuous(position = "top", breaks = 0:6) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    axis.text = element_text(size = 10),
    axis.title.x = element_text(size = 12, hjust = 0),
    axis.title.y = element_text(
      size = 12, hjust = 1,
      angle = 0, margin = margin(t = 0, r = 10, b = 0, l = 0)
    )
  ) +
  scale_colour_manual(
    breaks = paste(1:4),
    values = c("#D55E00", "#0072B2", "#009E73", "#CC79A7")
  ) +
  guides(alpha = FALSE)
```

## How does ARIMA() work?

```{r ARMAgridsearch2, echo=FALSE, message=FALSE, warning=FALSE, fig.asp=1, out.width="60%", fig.width=4, fig.height=4}
griddf |>
  ggplot(aes(x = q, y = p)) +
  geom_point(aes(alpha = 0.2), colour = "gray", size = 5, shape = 19) +
  geom_segment(aes(x = fromq, y = fromp, xend = q, yend = p, col = step),
    data = griddf |> filter(step == "2"),
    arrow = arrow(length = unit(0.15, "inches"), type = "open"),
    size = 1, lineend = "butt"
  ) +
  geom_point(aes(col = step),
    size = 5, shape = 19,
    data = griddf |> filter(step %in% c("1", "2"))
  ) +
  geom_point(data = filter(griddf |> filter(step %in% c("1", "2")), chosen), size = 12, shape = 21, stroke = 1.4) +
  scale_y_reverse(breaks = 0:6) +
  scale_x_continuous(position = "top", breaks = 0:6) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    axis.text = element_text(size = 10),
    axis.title.x = element_text(size = 12, hjust = 0),
    axis.title.y = element_text(
      size = 12, hjust = 1,
      angle = 0, margin = margin(t = 0, r = 10, b = 0, l = 0)
    )
  ) +
  scale_colour_manual(
    breaks = paste(1:4),
    values = c("#D55E00", "#0072B2", "#009E73", "#CC79A7")
  ) +
  guides(alpha = FALSE)
```

## How does ARIMA() work?

```{r ARMAgridsearch3, echo=FALSE, message=FALSE, warning=FALSE, fig.asp=1, out.width="60%", fig.width=4, fig.height=4}
griddf |>
  ggplot(aes(x = q, y = p)) +
  geom_point(aes(alpha = 0.2), colour = "gray", size = 5, shape = 19) +
  geom_segment(aes(x = fromq, y = fromp, xend = q, yend = p, col = step),
    data = griddf |> filter(step %in% "3"),
    arrow = arrow(length = unit(0.15, "inches"), type = "open"),
    size = 1, lineend = "butt"
  ) +
  geom_point(aes(col = step),
    size = 5, shape = 19,
    data = griddf |> filter(step %in% c("1", "2", "3"))
  ) +
  geom_point(
    data = griddf |> filter(step %in% c("1", "2", "3"), chosen),
    size = 12, shape = 21, stroke = 1.4
  ) +
  scale_y_reverse(breaks = 0:6) +
  scale_x_continuous(position = "top", breaks = 0:6) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    axis.text = element_text(size = 10),
    axis.title.x = element_text(size = 12, hjust = 0),
    axis.title.y = element_text(
      size = 12, hjust = 1,
      angle = 0, margin = margin(t = 0, r = 10, b = 0, l = 0)
    )
  ) +
  scale_colour_manual(
    breaks = paste(1:4),
    values = c("#D55E00", "#0072B2", "#009E73", "#CC79A7")
  ) +
  guides(alpha = FALSE)
```

## How does ARIMA() work?

```{r ARMAgridsearch4, echo=FALSE, message=FALSE, warning=FALSE, fig.asp=1, out.width="60%", fig.width=4, fig.height=4}
griddf |>
  ggplot(aes(x = q, y = p)) +
  geom_point(aes(alpha = 0.2), colour = "gray", size = 5, shape = 19) +
  geom_segment(aes(x = fromq, y = fromp, xend = q, yend = p, col = step),
    data = griddf |> filter(step %in% "4"),
    arrow = arrow(length = unit(0.15, "inches"), type = "open"),
    size = 1, lineend = "butt"
  ) +
  geom_point(aes(col = step),
    size = 5, shape = 19,
    data = griddf |> filter(step %in% c("1", "2", "3", "4"))
  ) +
  geom_point(
    data = griddf |> filter(step %in% c("1", "2", "3", "4"), chosen),
    size = 12, shape = 21, stroke = 1.4
  ) +
  scale_y_reverse(breaks = 0:6) +
  scale_x_continuous(position = "top", breaks = 0:6) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    axis.text = element_text(size = 10),
    axis.title.x = element_text(size = 12, hjust = 0),
    axis.title.y = element_text(
      size = 12, hjust = 1,
      angle = 0, margin = margin(t = 0, r = 10, b = 0, l = 0)
    )
  ) +
  scale_colour_manual(
    breaks = paste(1:4),
    values = c("#D55E00", "#0072B2", "#009E73", "#CC79A7")
  ) +
  guides(alpha = FALSE)
```

# Lab Session 16
## Lab Session 16

For the United States GDP data (from `global_economy`):

 * Fit a suitable ARIMA model for the logged data.
 * Produce forecasts of your fitted model. Do the forecasts look reasonable?

# Seasonal ARIMA models

## Seasonal ARIMA models
\fontsize{13}{15}\sf

| ARIMA | $~\underbrace{(p, d, q)}$ | $\underbrace{(P, D, Q)_{m}}$ |
| ----: | :-----------------------: | :--------------------------: |
|       | ${\uparrow}$              | ${\uparrow}$                 |
|       | Non-seasonal part         | Seasonal part of             |
|       | of the model              | of the model                 |

\vspace*{-0.4cm}

  * $m =$ number of observations per year.
  * $d$ first differences, $D$ seasonal differences
  * $p$ AR lags, $q$ MA lags
  * $P$ seasonal AR lags, $Q$ seasonal MA lags

###
Seasonal and non-seasonal terms combine multiplicatively

## Cortecosteroid drug sales

```{r, echo=TRUE}
h02 <- PBS |>
  filter(ATC2 == "H02") |>
  summarise(Cost = sum(Cost) / 1e6)
```

## Cortecosteroid drug sales

```{r, echo=TRUE}
h02 |> autoplot(
  Cost
)
```

## Cortecosteroid drug sales

```{r, echo=TRUE}
h02 |> autoplot(
  log(Cost)
)
```

## Cortecosteroid drug sales

```{r, echo=TRUE}
h02 |> autoplot(
  log(Cost) |> difference(12)
)
```

## Cortecosteroid drug sales

```{r, echo=TRUE}
h02 |> autoplot(
  log(Cost) |> difference(12) |> difference(1)
)
```

## Example: US electricity production
\fontsize{11}{12}\sf

```{r h02fit, echo=TRUE, fig.height=2.8}
h02 |>
  model(arima = ARIMA(log(Cost))) |>
  report()
```

## Example: US electricity production
\fontsize{11}{13}\sf

```{r h02fcst, echo=TRUE, fig.height=2.8}
h02 |>
  model(arima = ARIMA(log(Cost))) |>
  forecast(h = "3 years") |>
  autoplot(h02)
```

\vspace*{5cm}

## Cortecosteroid drug sales
\fontsize{9}{9}\sf

```{r h02tryharder, echo=TRUE, fig.height=3.6}
fit <- h02 |>
  model(best = ARIMA(log(Cost),
    stepwise = FALSE,
    approximation = FALSE,
    order_constraint = p + q + P + Q <= 9
  ))
report(fit)
```

## Cortecosteroid drug sales
\fontsize{11}{14}\sf

```{r h02f, echo=TRUE, fig.height=2.8}
fit |>
  forecast() |>
  autoplot(h02) +
  labs(y = "H02 Expenditure ($AUD)", x = "Year")
```

# Lab Session 17
## Lab Session 17

For the Australian tourism data (from `tourism`):

 * Fit a suitable ARIMA model for all data.
 * Produce forecasts of your fitted models.
 * Check the forecasts for the "Snowy Mountains" and "Melbourne" regions. Do they look reasonable?

# Forecast ensembles

## Forecast ensembles
\fontsize{10}{11}\sf

```{r trainall, echo=TRUE, dependson='tourism'}
train <- tourism |>
  filter(year(Quarter) <= 2014)
fit <- train |>
  model(
    ets = ETS(Trips),
    arima = ARIMA(Trips),
    snaive = SNAIVE(Trips)
  ) |>
  mutate(mixed = (ets + arima + snaive) / 3)
```

\fontsize{13}{14}\sf

 * Ensemble forecast `mixed` is a simple average of the three fitted models.
 *  `forecast()` will produce distributional forecasts taking into account the correlations between the forecast errors of the component models.

## Forecast ensembles
\fontsize{10}{11}\sf

```{r trainfc, dependson='trainall'}
#| fig-height: 3.6
fc <- fit |> forecast(h = "3 years")
fc |>
  filter(Region == "Snowy Mountains", Purpose == "Holiday") |>
  autoplot(tourism, level = NULL)
```

## Forecast ensembles
\fontsize{10}{11}\sf

```{r snowy-test-accuracy, dependson='trainfc'}
accuracy(fc, tourism) |>
  group_by(.model) |>
  summarise(
    RMSE = mean(RMSE),
    MAE = mean(MAE),
    MASE = mean(MASE)
  ) |>
  arrange(RMSE)
```

## Forecast ensembles
\fontsize{13}{15}\sf

\begin{alertblock}{}\bfseries Can we do better than equal weights?\end{alertblock}\pause\vspace*{-0.3cm}

 * Hard to find weights that improve forecast accuracy.
 * Known as the "forecast combination puzzle".
 * Solution:  FFORMA

\pause

\begin{block}{\fontsize{13}{14}\sf\textbf{FFORMA (Feature-based FORecast Model Averaging)}}
\begin{itemize}
\item Vector of time series features used to predict best weights.
\item A modification of xgboost is used.
\item Method came 2nd in the 2018 M4 international forecasting competition.
\item Main author: Pablo Montero-Manso (now Uni Sydney)
\item Not (yet) available for fable.
\end{itemize}
\end{block}
